Assessing United States county-level exposure for research on tropical cyclones and human health

### 2019-11-28 writing/ehp_introduction.tex
Continue work on text

Hurricanes and other tropical cyclones can severely impact human health in
\ac{US} communities, impacts that have been characterized based on fatality
data from post-disaster surveillance~\parencite{rappaport2000,
rappaport2014fatalities, rappaport2016fatalities, czajkowski2010fatal,
czajkowski2011, moore2012}.  For several types of climate-related
disasters---including heat waves, floods, and wildfires---research based on
post-disaster surveillance data has been richly supplemented by studies
analyzing trends in community-wide administrative or vital data, including time
series or case-crossover studies (e.g.,~\cite{anderson2010heat, son2012impact,
haikerwal2015impact, liu2017wildfire, milojevic2017mental}).  These studies
typically compare the rate of a health outcome in a community during a disaster
to rates during comparable non-disaster periods.  While such studies cannot
attribute specific cases (e.g., a specific death) to a disaster, they
can help characterize the community-wide change in health risk during disasters
and capture impacts that might be missed or underestimated in post-disaster
surveillance.  In many cases, these studies analyze multi-year, multi-community
data, allowing them to estimate average associations over many disasters and to
explore how a disaster's characteristics, or the characteristics of the
community it hits, modify associated health risks
(e.g.,~\cite{anderson2010heat, son2012impact, liu2017wildfire}).  Some studies
have begun to use this approach to study the health impacts of tropical
cyclones, including several studies of Hurricane Maria
(e.g.,~\cite{santos2018use, santos2018differential}), Hurricane Sandy
(e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane season
in Florida~\parencite{mckinney2011}.  However, to expand this approach to study
longer time periods for large sets of communities, researchers must be able to
assess exposure to tropical cyclones consistently and comparably across storms,
years, and communities.

The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
standard for Atlantic-basin tropical cyclone tracks.  It records the central
position of a tropical cyclone every six hours, as well as the storm's minimum
central pressure and maximum wind speed.  These data are openly available
through the \ac{HURDAT2}, a post-storm assessment that incorporates data from
several sources, including satellite data and, when available, aircraft
reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these data, it
is straightforward to measure whether a community was in the direct path, or
within a certain distance of the central track, of a storm, as some previous
epidemiological studies have done to assess exposure to a tropical storm
(e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}).

However, when epidemiological studies assess exposure to a tropical cyclone
using, as a proxy, how close the storm came to a community, they
may misclassify exposure~\parencite{grabich2015measuring}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and
waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
floods, and tornadoes can occur well away from the tropical cyclone's central
track~\parencite{rappaport2000, atallah2007, moore2012}.  For example, fatal
tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
tropical cyclone's center~\parencite{moore2012}.  Further, when studies use
distance from the storm's track to assess exposure, they often use the same
buffer constraints on each side of the storm track
(e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
cyclone tend to be distributed around its center non-symmetrically.  Extreme
winds are more common to the track's right, where counter-clockwise cyclonic
winds move in concert with the tropical cyclone's forward
motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
is often heaviest to the left of the track, especially when the tropical
cyclone interacts with other weather systems~\parencite{atallah2003,
atallah2007, zhu2013variations} or undergoes an extratropical
transition~\parencite{elsberry2002}.  The multi-hazard nature of tropical
cyclones therefore makes it hard to assess exposure based on how close the
storm's central track came to the community.  While other approaches have been
developed to incorporate storm hazards, particulary wind, into exposure
assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
there is not yet a standard approach, and when different studies use different
datasets or storm hazards when assessing storm exposure, it becomes difficult
to compare and aggregate findings from different studies.

To help epidemiological researchers assess exposure to tropical cyclones, we
developed an open-source dataset~\parencite{hurricaneexposure}.  These data
cover all counties in states in the eastern half of the \ac{US} over multiple
years and include five metrics characterizing exposure to tropical cyclones
(Table~\ref{tab:exposuremetrics}): (1)~closest distance the storm's central
track came to the county's center (a proxy measurement of storm exposure used
in some previous studies); (2)~maximum sustained wind speed at the county's
center over the course of the storm; (3)~cumulative rainfall in the county over
the course of the storm; (4)~flooding in the county
concurrent with the storm; and (5)~tornadoes in the county
concurrent with the storm.  We aggregated these data at the county level since
data for epidemiological studies are often available at county-level
aggregations (e.g., direct hurricane-related deaths~\parencite{czajkowski2011};
birth outcomes~\parencite{grabich2015, grabich2015measuring}; autism
prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
for, and respond to, tropical cyclones are often undertaken at the county
level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
explore and map these data and to link it with county-level health
data~\parencite{hurricaneexposuredata}

Here, we describe how we developed this dataset and explore how its
measurements compare to other data that could be used to characterize tropical
cyclone hazards for epidemiological research. Further, we expand
on previous research that investigated methods to measure exposure to tropical
cyclones for epidemiological research~\parencite{grabich2015measuring}.  The
previous study investigated differences in which communities were assessed as
exposed to tropical cyclones during the 2004 hurricane season in Florida,
depending on whether exposure assessment was based on distance to the storm's
track or used a metric incorporating storm-related winds within the county
during the storm.  They found important differences across methods of assessing
storm exposure, concluding that a study may be prone to bias from exposure
misclassification if distance to the storm track is used as a proxy measurement
for exposure assessment~\parencite{grabich2015measuring}.  Here we expand to
investigate this question across a large set of counties and storm seasons.
Further, we investigate patterns in exposure classification based on other
metrics of storm-related hazards---rainfall, flooding, and tornadoes---that are
important to characterize to improve understanding of the inland health impacts
of tropical cyclones~\parencite{czajkowski2011, moore2012}. These results can
help epidemiologists design studies and plan statistical analysis for
multi-year, multi-community studies that investigate changes in community-wide
rates of health outcomes during exposure to the hazards brought by tropical
cyclones, including counties inland from a storm's landfall.

>>>>

Hurricanes and other tropical cyclones can severely impact human health in
\ac{US} communities, as has been characterized in several long-term national
studies~\parencite{rappaport2000, rappaport2014fatalities,
rappaport2016fatalities, czajkowski2010fatal, czajkowski2011, moore2012}.
These studies characterize tropical cyclone health impacts using fatality data
aggregated in large part from disaster-related mortality ascertainment and
surveillance conducted by agencies like the \ac{US} Centers for Disease Control
and Prevention, the \ac{US} National Weather Service, and local vital
statistics departments. These fatality data are based on case-by-case
ascertainment: identifying and characterizing deaths for which there is a clear
link with a disaster, through death certificate coding or other indicators
\parencite{rocha2017medicolegal}.

For several types of climate-related disasters---including heat waves, floods,
and wildfires---such research has been richly supplemented by studies that
estimate community-wide excess mortality and morbidity associated with the
disaster (e.g.,~\cite{anderson2010heat, son2012impact, haikerwal2015impact,
liu2017wildfire, milojevic2017mental}).  These studies, which often use time
series or case-crossover study designs, typically compare the rate of a health
outcome in a community during a disaster to rates during comparable
non-disaster periods.  While such studies cannot attribute specific cases
(e.g., a specific death) to a disaster, they can quantify the community-wide
change in health risk during disasters and capture impacts that might be missed
or underestimated with traditional disaster-related mortality ascertainment and
surveillance.  In many cases, these studies analyze multi-year, multi-community
data, allowing them to estimate average associations over many disasters and to
explore how a disaster's characteristics, or the characteristics of the
community it hits, modify associated health risks
(e.g.,~\cite{anderson2010heat, son2012impact, liu2017wildfire}).  Some studies
have begun to use this approach to study the health impacts of tropical
cyclones, including several studies of Hurricane Maria
(e.g.,~\cite{santos2018use, santos2018differential}), Hurricane Sandy
(e.g.,~\cite{kim2016, mongin2017, swerdel2014}), and the 2004 hurricane season
in Florida~\parencite{mckinney2011}.  However, to expand this approach to
longer time periods and large sets of communities, researchers must be able to
assess exposure to tropical cyclones consistently and comparably across storms,
years, and communities.

The \ac{NHC} publishes a ``Best Tracks'' dataset that is considered the gold
standard for Atlantic-basin tropical cyclone tracks.  It records the central
position of a tropical cyclone every six hours, as well as the storm's minimum
central pressure and maximum sustained surface wind.  These data are openly
available through the \ac{HURDAT2}, a post-storm assessment that incorporates
data from several sources, including satellite data and, when available,
aircraft reconnaissance data~\parencite{landsea2013, jarvinen1988}.  With these
data, it is straightforward to measure whether a community was in a storm's
direct path or within a certain distance of its central track; some previous
epidemiological studies have done this to assess exposure to a tropical storm
(e.g.,~\cite{currie2013, kinney2008, caillouet2008increase}).

However, when epidemiological studies assess exposure to a tropical cyclone
using, as a proxy, how close the storm came to a community, they
may misclassify exposure~\parencite{grabich2015measuring}. While a number of
tropical cyclone hazards are strongly associated with distance from the
tropical cyclone's center (e.g., wind and, at the coast, storm surge and
waves~\parencite{rappaport2000, kruk2010}), other hazards like heavy rainfall,
floods, and tornadoes can occur well away from the tropical cyclone's central
track~\parencite{rappaport2000, atallah2007, moore2012}.  For example, fatal
tropical cyclone tornadoes, which were linked to over~300 deaths in the \ac{US}
between~1995 and~2009, most often occur~200\,--\,500~\si{\kilo\metre} from the
tropical cyclone's center~\parencite{moore2012}.  Further, when studies use
distance from the storm's track to assess exposure, they often use the same
buffer constraints on each side of the storm track
(e.g.,~\cite{kinney2008, currie2013}).  However, the forces of a tropical
cyclone tend to be distributed around its center non-symmetrically.  Extreme
winds are more common to the track's right, where counter-clockwise cyclonic
winds move in concert with the tropical cyclone's forward
motion~\parencite{halverson2015}, and almost all of the fatal tornadoes
associated with \ac{US} tropical cyclones between~1995 and~2009 occurred to the
right of the tropical cyclone's track~\parencite{moore2012}. Rain, conversely,
is often heaviest to the left of the track, especially when the tropical
cyclone interacts with other weather systems~\parencite{atallah2003,
atallah2007, zhu2013variations} or undergoes an extratropical
transition~\parencite{elsberry2002}.  The multi-hazard nature of tropical
cyclones therefore makes it hard to assess exposure based on how close the
storm's central track came to the community.  While other approaches have been
developed to incorporate storm hazards, particulary wind, into exposure
assessment (e.g., \cite{grabich2015measuring, zandbergen2009, czajkowski2011}),
there is not yet a standard approach, and when different studies use different
datasets or storm hazards when assessing storm exposure, it becomes difficult
to compare and aggregate findings.

To help epidemiological researchers assess exposure to tropical cyclones, we
developed an open-source dataset~\parencite{hurricaneexposure}, which we
present here.  These data cover all counties in states in the eastern half of
the \ac{US} over multiple years and include five metrics characterizing
exposure to tropical cyclones (Table~\ref{tab:exposuremetrics}): (1)~closest
distance the storm's central track came to the county's center (a proxy
measurement of storm exposure used in some previous studies); (2)~peak
sustained surface wind at the county's center over the course of the storm;
(3)~cumulative rainfall in the county over the course of the storm;
(4)~flooding in the county concurrent with the storm; and (5)~tornadoes in the
county concurrent with the storm.  We aggregated these data at the county level
since data for epidemiological studies are often available at county-level
aggregations (e.g., direct hurricane-related deaths~\parencite{czajkowski2011};
birth outcomes~\parencite{grabich2015, grabich2015measuring}; autism
prevalence~\parencite{kinney2008}) and since decisions and policies to prepare
for, and respond to, tropical cyclones are often undertaken at the county
level~\parencite{zandbergen2009, rappaport2000}.  We also wrote functions to
explore and map these data and to link it with county-level health
data~\parencite{hurricaneexposuredata}.

Here, we describe how we developed this dataset and explore how its
measurements compare to other data that could be used to characterize tropical
cyclone hazards for epidemiological research. Further, we expand on previous
research on methods to measure exposure to tropical cyclones for
epidemiological research~\parencite{grabich2015measuring}.  The previous study
investigated differences in which communities were assessed as exposed to
tropical cyclones during the 2004 hurricane season in Florida, depending on
whether exposure assessment was based on distance to the storm's track or a
metric incorporating storm-related winds within the county during the storm.
They found important differences across methods of assessing storm exposure,
concluding that a study may be prone to bias from exposure misclassification if
distance to the storm track is used as a proxy measurement for exposure
assessment~\parencite{grabich2015measuring}.  Here we expand to investigate
this question across a large set of counties and storm seasons.  Further, we
investigate patterns in exposure classification based on other metrics of
storm-related hazards---rainfall, flooding, and tornadoes---which are important
for inland health impacts of tropical cyclones~\parencite{czajkowski2011,
moore2012}. These results can help epidemiologists design studies and plan
statistical analysis for multi-year, multi-community studies that estimate
excess mortality and morbidity exposure to the hazards brought by tropical
cyclones.

### 2019-11-28 writing/ehp_methods.tex
Continue work on text

We aggregated data for all counties in the eastern half of the \ac{US}
(Figure~\ref{fig:hurrtracks}). We
included all tracked land-falling or near-land Atlantic-basin tropical cyclones
between~1988 and~2015 by covering all 136 storms in \ac{HURDAT2}
\parencite{landsea2013} that came within~250 \si{\kilo\metre} of at least one
eastern \ac{US} county in this period (the tracks of the study storms are shown in 
Figure~\ref{fig:hurrtracks}).

@@

To validate these precipitation estimates, we compared them with ground-based
observations in a subset of study counties. We selected nine sample counties
geographically spread across storm-prone regions of the eastern \ac{US} and for
which precipitation data were available from multiple ground-based stations
within the Global Historical Climatology Network throughout the study
period~\parencite{menne2012overview, rnoaa, countyweather}. These sample
counties were: Miami-Dade, FL; Harris County, TX; Mobile County, AL; Orleans
Parish, LA; Fulton County, GA; Charleston County, SC; Wake County, NC;
Baltimore County, MD; and Philadelphia County, PA. We summed daily
station-specific measurements from two days before to one day after each
storm's closest approach and then averaged these cumulative station-based
precipitation totals for each county to create a county-wide estimate of
cumulative storm-related precipitation based on ground-based monitoring. We
measured the rank correlation (Spearman's~$\rho$~\parencite{spearman1904proof})
between storm-specific cumulative precipitation estimates for the two data
sources (\ac{NLDAS-2} re-analysis data versus ground-based monitoring) within
each sample county.

@@

To create a dataset of county-level maximum sustained winds during historical
tropical cyclones, we modeled maximum sustained wind speeds at each county's
population mean center~\parencite{countycenters} using a wind speed model based
on results from Willoughby~\parencite{willoughby2006parametric}. For each
storm, we modeled the maximum ground-level sustained wind speed that the storm
generated at each county center for each~15-\si{\minute} increment throughout
the storm tracking period. We used a double exponential wind speed model with
inputs for the storm's forward speed, direction, and maximum wind
speed~\parencite{willoughby2006parametric, stormwindmodel}, incorporating the
asymmetry in wind speeds around the tropical cyclone center that results from
the storm's forward movement~\parencite{phadke2003modeling}. From these
15-\si{\minute} interval estimates, we identified the maximum sustained surface
wind speed in each county over the course of each storm.

To validate these modeled county-level wind speed estimates, we compared them
to the  post-season wind radii values, which have been routinely published
since 2004 \parencite{knaff2016using}. These wind radii characterize storm
size, estimating the maximum distance from the storm's center that winds of a
certain intensity extend. They give separate estimates for four quadrants of
the storm to capture asymmetry in the storm's wind field and are 
based on a post-season
re-analysis that incorporates available data (e.g., satellite data, aircraft
reconnaissance and ground-based data if available)
\parencite{knaff2016using}. Wind radii are estimated
for three thresholds of maximum surface wind speeds: 64,~50, and~34 \si{\knot}.
They therefore allow for the classification of counties into four categories of
maximum sustained wind speeds: $<$34~\si{\knot}; 34\,--\,49.9~\si{\knot};
50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}.

We interpolated the wind radii data to~15-\si{\minute} increments and
classified a county as exposed to winds in a given wind speed category if its
population mean center was within~85\% of the maximum radius for that wind
speed in that quadrant of the storm. The~85\% adjustment is based on previous
research that found that this ratio helps capture average wind extents within a
quadrant from these maximum wind radii \parencite{knaff2016using}.  We compared
these wind radii--based estimates of county-level maximum sustained surface
winds during a storm with the modeled wind speed estimates, comparing all study
storms since 2004 for which at least one study county experienced sustained
winds of~$\ge$34 \si{\knot} (based on the \ac{HURDAT2} wind radii). For each of
these storms, we calculated the percent of study counties that were classified
in the same wind speed category based on both data sources.

@@

To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)'s Storm Event
Database~\parencite{stormevents}. While this database has recorded storm data
since~1950, its coverage changed substantially in~1996 to cover a more types of
storm events~\parencite{stormevents}. We therefore only considered these
metrics of hurricane exposure for tropical cyclones in~1996 and later.

@@

In our open-source data, we provide continuous measurements of some of the
exposure metrics: closest distance of each storm to each county, maximum
sustained winds at the county center, and cumulative rainfall. However,
epidemiologic studies of tropical cyclones often use a binary exposure
classification (``exposed'' versus ``unexposed'') to assess storm-related
health risks, and so we explored patterns in storm exposure based on binary
clasifications of these exposure metrics.

@@

We characterized patterns in county-level exposure in the eastern \ac{US} for
exposure assessment based on each of the measured tropical cyclone hazards
(cumulative precipitation, maximum sustained winds, flooding, and tornadoes).
Depending on available exposure data, this assessment included some or all of
the period from~1988 to~2015 (Table 1). For each binary metric of exposure to a
tropical cyclone hazard, we first summed the total number of county-level
exposures over available years and mapped patterns in
these exposures.

Next, we investigated how well exposure assessment agreed across 
these metrics. For each storm we investigated the degree to which the 
set of counties assessed as exposed based on one metric overlapped with the set
assessed as exposed based on each other metric, including a metric that used
distance of a county from the storm track as a proxy for exposure to storm hazards. 
We calculated the within-storm
Jaccard index ($J_s$)~\parencite{jaccard1901distribution, jaccard1908nouvelles} between
each pair of exposure metrics. This measures similarity
between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical cyclone~$s$ as the
proportion of counties in which both of the metrics classify the county as
exposed out of all counties classified as exposed by at least one of the
metrics:

>>>>

We aggregated these data for all counties in the eastern half of the \ac{US}
(Figure~\ref{fig:hurrtracks}).  We included all tropical cyclones in
\ac{HURDAT2} \parencite{landsea2013} that came within~250 \si{\kilo\metre} of
at least one eastern \ac{US} county between 1988 and 2015.  This ensured we
were including all land-falling or near-land Atlantic basin tropical cyclones
in this period (the tracks of the study storms are shown in
Figure~\ref{fig:hurrtracks}).

@@

To validate these precipitation estimates, we compared them with ground-based
observations in a subset of study counties. We selected nine sample counties
geographically spread across storm-prone regions of the eastern \ac{US} and for
which precipitation data were available from multiple ground-based stations
within the Global Historical Climatology Network throughout the study
period~\parencite{menne2012overview, rnoaa, countyweather}. These sample
counties were: Miami-Dade, FL; Harris County, TX; Mobile County, AL; Orleans
Parish, LA; Fulton County, GA; Charleston County, SC; Wake County, NC;
Baltimore County, MD; and Philadelphia County, PA. We summed daily
station-specific measurements from two days before to one day after each
storm's closest approach and then averaged these cumulative station-based
precipitation totals for each county to create a county-wide estimate of
cumulative storm-related precipitation based on ground-based monitoring. We
measured the rank correlation (Spearman's~$\rho$;~\textcite{spearman1904proof})
between storm-specific cumulative precipitation estimates for the two data
sources (\ac{NLDAS-2} re-analysis data versus ground-based monitoring) within
each sample county.

@@

We created a dataset of county-level peak sustained surface wind during each
storm. To do this, we first modeled each storm's wind field to each county's
population mean center~\parencite{countycenters} at 15-minute intervals
throughout the course of the storm. We used a double exponential wind model
based on results from Willoughby~\parencite{willoughby2006parametric} to model
1-minute surface wind (10 meters above ground) at each county center, based on
inputs of the storm's forward speed, direction, and maximum wind
speed~\parencite{stormwindmodel}. Our model incorporated the asymmetry in wind
speeds around the tropical cyclone center that results from the storm's forward
movement~\parencite{phadke2003modeling}.  From these 15-\si{\minute} interval
estimates, we identified the peak sustained surface wind in each county over
the course of each storm.

To validate these modeled county-level peak wind estimates, we compared them
to the  post-season wind radii values, which have been routinely published
since 2004 \parencite{knaff2016using}. These wind radii estimate the maximum
distance from the storm's center that winds of a certain intensity extend. They
give separate estimates for four quadrants of the storm to capture asymmetry in
the storm's wind field and are based on a post-season re-analysis that
incorporates available data (e.g., satellite data, aircraft reconnaissance and
ground-based data if available) \parencite{knaff2016using}. Wind radii are
estimated for three thresholds of peak sustained surface wind: 64,~50, and~34
kt.  They therefore allow for the classification of counties into four
categories of peak sustained surface wind: $<$34~kt; 34\,--\,49.9~kt;
50\,--\,63.9~kt; $\ge$64~kt.

We interpolated the wind radii data to~15-minute increments and
classified a county as exposed to winds in a given wind speed category if its
population mean center was within~85\% of the maximum radius for that wind
speed in that quadrant of the storm. The~85\% adjustment is based on previous
research that found that this ratio helps capture average wind extents within a
quadrant from these maximum wind radii \parencite{knaff2016using}.  We compared
these wind radii--based estimates of county-level peak sustained surface wind
during a storm with the modeled wind estimates, comparing all study storms
since 2004 for which at least one study county experienced peak sustained winds
of~$\ge$34 kt (based on the \ac{HURDAT2} wind radii). For each of these storms,
we calculated the percent of study counties that were classified in the same
wind category based on both data sources.

@@

To identify flood- and tornado-based tropical cyclone exposures in \ac{US}
counties, we matched storm tracks with event listings from the National Oceanic
and Atmospheric Administration (NOAA)'s Storm Event
Database~\parencite{stormevents}. While this database has recorded storm data
since~1950, its coverage changed substantially in~1996 to cover more types of
storm events~\parencite{stormevents}. We therefore only considered these
metrics of hurricane exposure for tropical cyclones in~1996 and later.

@@

In our open-source data, we provide continuous measurements of some of the
exposure metrics: closest distance of each storm to each county, peak sustained
surface wind at the county center, and cumulative rainfall. However,
epidemiologic studies of tropical cyclones often use a binary exposure
classification (``exposed'' versus ``unexposed'') to assess storm-related
health risks (e.g., \textcite{grabich2015, mckinney2011,
caillouet2008increase}), and so we explored patterns in storm exposure based on
binary clasifications of these exposure metrics.

@@

We characterized patterns in county-level exposure in the eastern \ac{US} for
exposure assessment based on each of the measured tropical cyclone hazards
(cumulative precipitation, peak sustained winds, flooding, and tornadoes).
Depending on available exposure data, this assessment included some or all of
the period from~1988 to~2015 (Table 1). For each binary metric of exposure to a
tropical cyclone hazard, we first summed the total number of county-level
exposures over available years and mapped patterns in
these exposures.

Finally, we investigated how well exposure assessment agreed across 
these metrics. For each storm we investigated the degree to which the 
set of counties assessed as exposed based on one metric overlapped with the set
assessed as exposed based on each other metric, including a metric that used
distance of a county from the storm track as a proxy for exposure to storm hazards. 
We calculated the within-storm
Jaccard index ($J_s$)~\parencite{jaccard1901distribution, jaccard1908nouvelles} between
each pair of exposure metrics. This measures similarity
between two metrics ($X_{1,s}$ and~$X_{2,s}$) for tropical cyclone~$s$ as the
proportion of counties in which both of the metrics classify the county as
exposed out of all counties classified as exposed by at least one of the
metrics:

### 2019-11-28 writing/ehp_results.tex
Continue work on text

We created and published this tropical cyclone exposure data as an R
package~\parencite{hurricaneexposuredata}.  The package size exceeded the
recommended maximum size for the standard repository for publishing R packages
(\ac{CRAN}). Therefore, we used the \textit{drat} framework to set up our own
repository to host the package~\parencite{anderson2017hosting}. These data
include continuous county-level measurements for the closest distance of each
storm, cumulative rainfall, and maximum sustained surface wind speed. These
continuous metrics can be used for classifying counties as exposed or
unexposed---using thresholds selected by the user---or can be extracted
directly as continuous metrics. The dataset also include binary data on flood
and tornado events associated with each storm in each county. To accompany
these data, we published an associated R package on \ac{CRAN} with software
tools to explore and map the data and to integrate it with human health
datasets~\parencite{hurricaneexposure}.

@@

For maximum sustained surface wind estimates, we found that the primary wind
exposure metric in the open-source data generally agreed well with the wind
radii reported in \ac{HURDAT2}. For most storms, $\ge$95\% of counties were
assigned the same category of wind speed ($<$34~\si{\knot};
34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot}; $\ge$64~\si{\knot}) by both
data sources (Figure~\ref{fig:windcomparison}).  Disagreement was limited to a
few storms (e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two
storms, the modeled wind speed values used in the open-source data somewhat
overestimated the extent of the storm's most severe winds at landfall but then
underestimated the extent of 34\,--\,49.9 \si{\knot} storm winds further inland
(Figure~S2). For epidemiology researchers who would like to conduct sensitivity
analysis between these two sets of wind data, we have included estimates from
the \ac{HURDAT2} wind radii as a secondary measure of wind speeds in the
open-source dataset \parencite{hurricaneexposuredata}.

@@

When we calculated and mapped the average number of exposures per decade in
each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
strong geographical patterns were clear. Tropical cyclone wind exposure had a
strong coastal pattern, with almost all exposures in counties within about~200
\si{\kilo\metre} (124~mi) of the coastline. While tropical cyclone rain
exposure was also more frequent in coastal areas compared to inland areas,
there were also inland rain exposures in counties that were never classified as
exposed based on winds. Flood-based exposures were frequently in the
Mid-Atlantic region, with a pattern that skewed north compared with other
exposures. Rain and, to some extent, flood exposures were characterized by a
pattern defined by the Appalachian Mountains, with fewer exposures west of the
mountain range than to the east. Almost all tornado-based exposures were in
coastal states, with many in Florida, and almost none north of Maryland.
Patterns were similar when analysis was restricted to years with exposure data
for all four hazards available (1996\,--\,2011; Figure~S3).

@@

Next, we measured the agreement between exposure classifications for each pair
of hazards, as well as with a proxy measurement based on
distance from the storm's track. We found that these exposure classifications
typically did not agree strongly between pairs of metrics---the set of counties
identified as ``exposed'' based on one metric often overlapped little with the set
identified as ``exposed'' by another metric, as storms frequently brought
different hazards to different locations.

@@

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (the period for which data for all five metrics were
available) for which~100 or more counties were exposed based on at least one
metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
for one tropical cyclone, and each box in that row shows the Jaccard
coefficient for a pair of metrics.  For almost all tropical cyclones, Jaccard
coefficients were~$\le0.6$ between the distance-based metric and any of the
four hazard-based metrics (i.e., out of the counties assessed as exposed by at
least one of the two metrics in the pair,~60\si{\percent} or fewer were
assessed the same under both metrics). For many tropical cyclones, Jaccard
coefficients were~$\le0.5$, indicating at best moderate agreement.  Among the
four hazard-based metrics, the tornado-based metric had universally poor
agreement with other metrics in county-level classification across the tropical
cyclones considered.  Agreement between other pairs of metrics were better, but
still typically at best moderate (Jaccard coefficients usually $\le0.5$).

There were, however, exceptions\,---\,tropical cyclones in which exposure
assessment agreed well across several of the metrics considered.  For Floyd
in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
agreed moderately to well for all pairs of exposure metrics except those
involving the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
in~2003), there was moderate to good agreement for all pairwise combinations of
distance, rain, and wind, but poor agreement for other combinations of metrics,
while for another set of storms (e.g., Matthew in~2004 and Katrina in~2005),
there was moderate to good agreement between distance and rain.  However, these
sets of tropical cyclones represented the minority of tropical cyclones
considered; for most tropical cyclones, there was low to moderate overlap in
the counties determined to be exposed to the tropical cyclones between most
pairings of metrics.

>>>>

We created and published this tropical cyclone exposure data as an R
package~\parencite{hurricaneexposuredata}.  The package size exceeded the
recommended maximum size for \ac{CRAN}, the standard repository for publishing
R packages. Therefore, we used the \textit{drat} framework to set up our own
repository to host the package~\parencite{anderson2017hosting}. These data
include continuous county-level measurements for the closest distance of each
storm, cumulative rainfall, and peak sustained surface wind. These continuous
metrics can be used for classifying counties as exposed or unexposed---using
thresholds selected by the user---or can be extracted directly as continuous
metrics. The dataset also include binary data on flood and tornado events
associated with each storm in each county. To accompany these data, we
published an additional R package with software tools to explore and map the
data and to integrate it with human health
datasets~\parencite{hurricaneexposure}.

@@

For peak sustained surface wind estimates, we found that the primary wind
exposure metric in the open-source data generally agreed well with the wind
radii reported in \ac{HURDAT2}. For most storms, $\ge$95\% of counties were
assigned the same category of wind speed ($<$34~kt; 34\,--\,49.9~kt;
50\,--\,63.9~kt; $\ge$64~kt) by both data sources
(Figure~\ref{fig:windcomparison}).  Disagreement was limited to a few storms
(e.g., Hurricanes Sandy in 2012 and Ike in 2008). For these two storms, the
modeled wind speed values used in the open-source data somewhat overestimated
the extent of the storm's most severe winds at landfall but then underestimated
the extent of 34\,--\,49.9 kt storm winds further inland (Figure~S2). For
epidemiology researchers who would like to conduct sensitivity analysis between
these two sets of wind data, we have included estimates from the \ac{HURDAT2}
wind radii as a secondary measure of county-level peak sustained wind in the
open-source dataset \parencite{hurricaneexposuredata}.

@@

When we calculated and mapped the average number of exposures per decade in
each county for single-hazard exposures (Figure~\ref{fig:averageexposure}),
strong geographical patterns were clear. Peak sustained wind exposure had a
strong coastal pattern, with almost all exposures in counties within about~200
\si{\kilo\metre} (124~mi) of the coastline. While tropical cyclone rain
exposure was also more frequent in coastal areas compared to inland areas,
there were also inland rain exposures in counties that were never classified as
exposed based on wind. Flood-based exposures were frequently in the
Mid-Atlantic region, with a pattern that skewed north compared with other
exposures. Rain and, to some extent, flood exposures were characterized by a
pattern defined by the Appalachian Mountains, with fewer exposures west of the
mountain range than to the east. Almost all tornado-based exposures were in
coastal states, with many in Florida, and almost none north of Maryland.
Patterns were similar when analysis was restricted to years with exposure data
for all four hazards available (1996\,--\,2011; Figure~S3).

@@

Finally, we measured the agreement between exposure classifications for each pair
of hazards, as well as with a proxy measurement based on
distance from the storm's track. We found that these exposure classifications
typically did not agree strongly between pairs of metrics---the set of counties
identified as ``exposed'' based on one metric often overlapped little with the set
identified as ``exposed'' by another metric, as storms frequently brought
different hazards to different locations.

@@

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (when data for all five metrics were
available) for which~100 or more counties were exposed based on at least one
metric (Figure~\ref{fig:jaccard}). In this figure, each row provides results
for one tropical cyclone, and each box in that row shows the Jaccard
coefficient for a pair of metrics.  For almost all tropical cyclones, Jaccard
coefficients were~$\le0.6$ between the distance-based metric and any of the
four hazard-based metrics (i.e., out of the counties assessed as exposed by at
least one of the two metrics in the pair,~60\si{\percent} or fewer were
assessed the same under both metrics). For many tropical cyclones, Jaccard
coefficients were~$\le0.5$, indicating at best moderate agreement.  Among the
four hazard-based metrics, the tornado-based metric had universally poor
agreement with other metrics in county-level classification across the tropical
cyclones considered.  Agreement between other pairs of metrics were better, but
still typically at best moderate (Jaccard coefficients usually $\le0.5$).

There were, however, exceptions\,---\,tropical cyclones in which exposure
assessment agreed well across several of the metrics considered.  For Floyd
in~1999 (Figure~S4) and Irene in~2011, for example, county-level classification
agreed moderately to well for all pairs of exposure metrics except those
involving the tornado-based metric (Figure~\ref{fig:jaccard}). For another set
of tropical cyclones (e.g., Ernesto in~2006, and Bertha in~1996, Isabel
in~2003), there was moderate to good agreement for all pairwise combinations of
distance, rain, and wind, but poor agreement for other combinations of metrics,
while for another set of storms (e.g., Matthew in~2004 and Katrina in~2005),
there was moderate to good agreement between distance and rain.
